---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# melting5jars

An R package wrapper for [MELTING 5](http://www.ebi.ac.uk/biomodels/tools/melting/) jars.

The 'MELTING 5' 'Java' archive file is included along with the model data directory to facilitate accessing  the library computations from within an R session.

## Installation

You can install `melting5jars` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("hrbrmstr/melting5jars")
```

## Example

This is an example made for the [SO question](https://stackoverflow.com/questions/48128038/r-wrapper-for-a-java-method-in-a-jar-using-rjava/48129909#48129909) that caused the creation of the package:

It is designed to (lightly) mimic the functionality of `melting/Main.java`

Here's `melting/Main.java`:

```{java melting_main, eval=FALSE}
package melting;

import java.text.NumberFormat;

import melting.configuration.OptionManagement;
import melting.configuration.RegisterMethods;
import melting.methodInterfaces.MeltingComputationMethod;
import melting.nearestNeighborModel.NearestNeighborMode;

/**
 * The Melting main class which contains the public static void main(String[] args) method.
 */
public class Main {

    // private static methods

    /**
     * Compute the entropy, enthalpy and the melting temperature and display the results. 
     * @param args : contains the options entered by the user.
     * @param OptionManagement optionManager : the OptionManegement which allows to manage
     * the different options entered by the user.
     */
    private static ThermoResult runMelting(String [] args, OptionManagement optionManager){
        try {
                        ThermoResult results = 
                                        getMeltingResults(args, optionManager);
            displaysMeltingResults(results);
                        return results;

        } catch (Exception e) {
            OptionManagement.logError(e.getMessage());
                        return null;
        }
    }

        /**
         * Compute the entropy, enthalpy and melting temperature, and return 
         * these results.
         * @param args options (entered by the user) that determine the
         *             sequence, hybridization type and other features of the
         *             environment.
         * @param optionManager the {@link 
         *                           melting.configuration.OptionManagement 
         *                           <code>OptionManagement</code>} which
         *                      allows the program to manage the different
         *                      options entered by the user.  
         * @return The results of the Melting computation.
         */
        public static ThermoResult getMeltingResults(String[] args,
                                                OptionManagement optionManager)
        {
            NumberFormat format = NumberFormat.getInstance();
            format.setMaximumFractionDigits(2);

            // Set up the environment from the supplied arguments and get the 
            // results.
            Environment environment = optionManager.createEnvironment(args);
            RegisterMethods register = new RegisterMethods();
            MeltingComputationMethod calculMethod = 
                register.getMeltingComputationMethod(environment.getOptions());
            ThermoResult results = calculMethod.computesThermodynamics();
            results.setCalculMethod(calculMethod);
            environment.setResult(results);

            // Apply corrections to the results.
            results = calculMethod.getRegister().
                                   computeOtherMeltingCorrections(environment);
            environment.setResult(results);
            return environment.getResult();
        }

    /**
     * displays the results of Melting : the computed enthalpy and entropy (in cal/mol and J/mol), and the computed 
     * melting temperature (in degrees).
     * @param results : the ThermoResult containing the computed enthalpy, entropy and
     * melting temperature
     * @param MeltingComputationMethod calculMethod : the melting computation method (Approximative or nearest neighbor computation)
     */
    private static void displaysMeltingResults(ThermoResult results)
        {
        NumberFormat format = NumberFormat.getInstance(); 
        format.setMaximumFractionDigits(2);
                MeltingComputationMethod calculMethod = 
                                                     results.getCalculMethod();

        double enthalpy = results.getEnthalpy();
        double entropy = results.getEntropy();

        OptionManagement.logInfo("\n The MELTING results are : ");
        if (calculMethod instanceof NearestNeighborMode){
            OptionManagement.logInfo("Enthalpy : " + format.format(enthalpy) + " cal/mol ( " + format.format(results.getEnergyValueInJ(enthalpy)) + " J /mol)");
            OptionManagement.logInfo("Entropy : " + format.format(entropy) + " cal/mol-K ( " + format.format(results.getEnergyValueInJ(entropy)) + " J /mol-K)");
        }
        OptionManagement.logInfo("Melting temperature : " + format.format(results.getTm()) + " degrees C.\n");
    }

    // public static main method

    /**
     * @param args : contains the options entered by the user.
     */
    public static void main(String[] args) {

        OptionManagement optionManager = new OptionManagement();

        if (args.length == 0){
            optionManager.initialiseLogger();
            optionManager.readMeltingHelp();
        }
        else if (optionManager.isMeltingInformationOption(args)){
            try {
                optionManager.readOptions(args);

            } catch (Exception e) {
                OptionManagement.logError(e.getMessage());
            }
        }
        else {
            runMelting(args, optionManager);
        }
    }
}
```

and, here's a light wrapper for it:

```{r wrapper}
get_melting_results <- function(opts = c()) {

  stopifnot(length(opts) > 2) # a sanity check that could be improved

  Sys.setenv("NN_PATH"=system.file("extdata", "Data", package="melting5jars"))

  require(melting5jars)

  melting <- new(J("melting.Main"))
  optionManager <- new(J("melting.configuration.OptionManagement"))

  results <- melting$getMeltingResults(opts, optionManager)

  calculMethod <- results$getCalculMethod()

  enthalpy_cal <- results$getEnthalpy()
  entropy_cal <- results$getEntropy()

  enthalpy_J <- entropy_J <- NULL

  if (.jinstanceof(calculMethod, J("melting.nearestNeighborModel.NearestNeighborMode"))) {
    enthalpy_J <- results$getEnergyValueInJ(enthalpy_cal)
    entropy_J <- results$getEnergyValueInJ(entropy_cal)
  }

  melting_temp_C <- results$getTm()

  list(
    enthalpy_cal = enthalpy_cal,
    entropy_cal = entropy_cal,
    enthalpy_J = enthalpy_J,
    entropy_J = entropy_J,
    melting_temp_C = melting_temp_C
  ) -> out

  class(out) <- c("melting_res")

  out

}

print.melting_res <- function(x, ...) {

  cat(
    "The MELTING results are:\n\n",
    "  - Enthalpy: ", prettyNum(x$enthalpy_cal), " cal/mol",
    {if (!is.null(x$enthalpy_J)) paste0(" (", prettyNum(x$enthalpy_J), " J /mol)", collapse="") else ""}, "\n",
    "  - Entropy: ", prettyNum(x$entropy_cal), " cal/mol-K",
    {if (!is.null(x$entropy_J)) paste0(" (", prettyNum(x$entropy_J), " J /mol-K)", collapse="") else ""}, "\n",
    "  - Meltng temperature: ", prettyNum(x$melting_temp_C), " degress C\n",
    sep=""
  )

}


Sodium <- 0.05

opts <- c(
  "-S", "GTCGTATCCAGTGCAGGGTCCGAGGTATTCGCACTGGATACGACTTCCAC",
  "-H", "dnadna",
  "-P", 5e-8,
  "-E", paste("Na=", Sodium, sep = "")
)

res <- get_melting_results(opts)

res

str(res)
```
